The Hot Hand Fallacy in Call of Duty

### 2022-08-22 paper/paper.tex
one big self review

\title{The Hot Hand in Call of Duty Search and Destroy}

@@

\begin{abstract}
Our research investigates patterns in round win percentages in
professional Search and Destroy (SnD) matches of the popular
first-person shooter game Call of Duty (CoD). First, we find evidence
supprting the hypothesis that round win probability can be modeled as a
constant across the series, although not at the naive 50\%. Second, we
examine post-streak round win probability, given the series outcome. We
find no evidence that post-streak win rate is significantly different
than expected for combinations of streak length, series length, and
series winner, suggesting that the perception of momentum at the team
level is just that, perception.
\end{abstract}

@@

Call of Duty (CoD), first released in 2003, is one of the most popular
first-person shooter (FPS) video game franchises of all-time. The most
popular mode in the competitive scene is ``Search and Destroy''
(SnD).\footnote{SnD bears resemblance to ``Bomb Defusal'' in
  Counter-Strike and ``Plant/Defuse'' in Valorant, two other FPS games
  played in more popular professional leagues.} SnD is a one-sided game
mode in which one team, the offensive side, tries to destroy one of two
designated bomb sites on the map.

In professional CoD SnD, the two teams take turns playing offense and
defense every round. They must win six rounds to win the
series.\footnote{A maximum of 11 even rounds can be played. There is no
  ``sudden death'' or ``win by two'' rule like there are for SnD
  equivalent in professional Counter-Strike and Valorant matches.} A
round can end in one of five ways:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  One team eliminates all members of the other team prior to a bomb
  plant. (Eliminating team wins.)
\item
  The offensive team eliminates all members of the defensive team after
  a bomb plant.\footnote{\begin{itemize}
    \tightlist
    \item
      The bomb can be picked up by any member of the offensive team.
    \item
      The bomb carrier is not obstructed at all by carrying the bomb
      (i.e.~movement is the same, weapon usage is the same).
    \item
      The defense does not get any visual indication for who is carrying
      the bomb.
    \item
      A bomb plant takes five seconds. The timer resets if the player
      stops planting site prior to completing it.
    \item
      A bomb defuse takes seven seconds. The timer resets if the player
      ``drops'' the bomb.
    \item
      The bomb takes 45 seconds to defuse after being planted.
    \end{itemize}} (Offense wins.)
\item
  The defensive team defuses the bomb after a bomb plant.\footnote{Often
    the defensive team will try to eliminate all team members prior to
    making the defuse, but in some cases, they may try to ``ninja''
    defuse.} (Defense wins.)
\item
  The offensive team does not make a plant by the time the round timer
  ends. (Defense wins.)
\end{enumerate}

We adopt the terminology ``series'' to refer to what CoD SnD players
typically call a ``match'', so as to emulate the terminology of playoff
series in professional leagues like the National Basketball Association,
National Hockey League, and Major League Baseball. A ``game'' or a
``match'' in such leagues is analogous to a ``round'' of CoD SnD.
Further, it is analogous to a trial in statistical analysis of Bernoulli
events.

SnD is not the only game type played in competitive CoD. Teams play in a
head-to-head, best-of-five (first-to-three wins) format, where SnD is
always played as the second and fifth game mode (if a fifth variant is
needed). The best-of-five across game types could also be called a
``series'', but since we analyze only the SnD game types, we refer to
SnD games as series (of rounds).

Anecdotally, most observers swear by the existence of momentum in CoD
SnD series, to the extent that vernacular has been developed to describe
such phenomenon. People say that teams ``ice up'' when the team wins
several consecutive rounds to come back to win a series, often
culminating with a series-clinching round win.\footnote{This vernacular
  is not specific to CoD SnD in specific, and is not only used in the
  contexts of streaks.} Further, viewers have come to embrace the
``5-3'' phenomenon, where teams win three consecutive rounds after
facing a 5-3 deficit to win 6-5. There is even a term for the rare 0-5
comeback---a ``full sail''.

@@

CoD has roughly gone through three eras of professional gaming: (1)
Major League Gaming (MLG) tournaments prior to 2016; (2) the CoD World
League (CWL), initiated in 2016; and (3) the 12-franchise CoD League
(CDL), operating since 2020. The CDL has completed three year-long
``seasons'' as of August 2022.\footnote{CoD is fairly unique compared to
  other esports in that it runs on an annual lifecycle (released coming
  in the late fall), where a new game is published every year under the
  same title. Each new game bears resemblance to past ones, often
  introducing relatively small variations (``improvements'') to
  graphics, game modes, and other facets of gameplay. During the CDL
  era, the games released have been Modern Warfare (2020), Cold War
  (2021) and Vanguard (2022).}

The data set consists of all SnD matches played in major tournaments and
qualifiers during the CDL era, totaling 7,792 rounds across 852
series.\footnote{288 of the series occur in major tournaments, which are
  considered to be more ``competitive'' than the qualifiers since they
  are played in person (COVID-permitting), whereas the qualifiers are
  played online.} Data was collected in spreadsheets by community member
``IOUTurtle''.\footnote{Raw data: \url{https://linktr.ee/CDLArchive}.
  Author: \url{https://twitter.com/IOUTurtle}. Processed data can be
  found at
  \url{https://github.com/tonyelhabr/fps_round_games/blob/master/data/cod_rounds.csv}.}

The observed offensive round win percentage across all rounds is
\(\tau_O\) = 47.8\%.\footnote{Offensive round win percentage has been
  nearly constant across the three games during the CDL era: 1. 47.2\%
  in MW (2020) 2. 47.9\% in Cold War (2021) 3. 48.1\% in Vanguard (2022)}
Table \ref{tbl:cod-o-win-prop-by-series-state} shows round win
percentages by series ``state'' (i.e.~the number of round wins by each
team prior to an upcoming round). Offensive round win rate is not quite
constant, although never veers more than 10\% from this overall average.

@@

On the extremes, 6-0 sweeps make up 4.7\% of all series, while 6-5
series make up 23.7 of series.

At tournaments, higher-seeded (``better'') teams choose whether they
want to start on offense in the SnD series played as the second game or
the fifth game, and their opponent is assigned to play offense in the
first round of the SnD slot not chosen.\footnote{``Hardpoint'' has been
  played as the first and fourth games in a matchup; the variant for the
  third game was Domination in the 2020 season, and Control in the 2021
  and 2022 seasons.}\footnote{Anecdotally, the higher-seeded teams tend
  to choose to play defense in the first round of the SnD series played
  as the second game, although this choice is not consistent across
  teams, or with the same team over time.} ({``Competitive Settings
\textbar{} Call of Duty League,''} n.d.) With online qualifiers, teams
all play each other twice throughout the season. Each team gets to play
the higher-seeded role once in their matchups, despite their win-loss
records. Overall, we might say that bias due to stronger teams playing
defense (slightly advantageous) more often is very minimal.

@@

There have been a handful of studies of the distribution of games played
in a series of a professional sport. Most assume a constant probability
\(\phi\) of a given team winning a game in the series, regardless of the
series state. Mosteller (1952) observed that the American League had
dominated the National League in Major League Baseball's (MLB) World
Series matchups, implying that games should not be modeled with a
constant (null) \(\phi_0 = 0.5\). Mosteller proposed three approaches
for identifying the optimal constant probability value of the stronger
team in the World Series, finding \(\phi \approx 0.65\) in each case:
(1) solving for \(\phi\) from the observed average number of games won
by the loser of the series, which he called the ``method of moments''
approach; (2) maximizing the likelihood that the sample would have been
drawn from a population in which the probability of a team winning a
game is constant across the series (i.e.~maximum likelihood), and (3)
minimizing the chi-square goodness of fit statistic \(\chi^2\) for
\(\phi\).

Chance (2020) re-examines the constant probability notion in Major
League Baseball's (MLB) World Series (1923--2018), the National
Basketball Association's (NBA) Finals (1951--2018), and the National
Hockey League's (NHL) Stanley Cup (1939--2018). Chance finds strong
evidence against the null hypothesis of \(\phi_0 = 0.5\) in the MLB and
NHL championship series when applying Mosteller's first and second
methods. Chance goes on to outline a conditional probability framework
(likelihood of winning a game given the series state) which can exactly
explain the distribution of the number of games played.

Momentum, one of most discussed topics in sports analytics, goes
hand-in-hand with a discussion of the nature of series
outcomes.\footnote{We often use use ``streaks'' and momentum
  interchangeably, but as (Steeger, Dulin, and Gonzalez 2021) note,
  momentum implies dependence between events, whereas streaking does
  not.} Two opposing fallacies are observed in the context of momentum:
the ``gambler's fallacy'' (negative recency) and the ``hot hand
fallacy'' (positive recency). Per Ayton et al. (2004), negative recency
is ``the belief that, for random events, runs of a particular outcome
\ldots{} will be balanced by a tendency for the opposite outcome'',
while positive recency is the expectation of observing future results
that match recent results.

Studying both player streaks and team streaks in basketball, in both
observational and controlled settings. Gilovich et al. (1985),
henceforth GVT, do not find evidence for the hot hand phenomenon.
However, Miller and Sanjurjo (2018), henceforth MS, provide a framework
for quantifying streak selection bias, which effectively works in the
manner posited by the gambler's fallacy. Specifically, MS say that a
``bias exists in a common measure of the conditional dependence of
present outcomes on streaks of past outcomes in sequential data''
implying that, under i.i.d. conditions, ``the proportion of successes
among the outcomes that immediately follow a streak of consecutive
successes is expected to be strictly less than the underlying
(conditional) probability of success''. When applying streak selection
bias to GVT's data, MS come to the opposite conclusions as GVT.

Other research has borrowed a common approach from the field of quality
control, looking at unlikely sequences of events with the Wald-Wolfowitz
runs test (Peel and Clauset 2015), (Steeger, Dulin, and Gonzalez 2021).
Peel and Clauset find no evidence for unlikely sequences of scoring
events in the NHL, College Football (CFB), and National Football League
(NFL), but do in the NBA. As a check on their entropy approach to
momentum identification, Steeger et al.~found three NHL teams who
sequence of wins in the 2018-2019 regular season violated the
Wald-Wolfowitz null hypothesis, although these teams did not match the
teams identified with their primary approach.

@@

Despite the plethora of existing research on games played in a series
and momentum in sports, these topics have yet to be investigated
heavily, if at all, in esports. Work has been done to examine
intra-round win probability in other FPS titles such as Counter-Strike
(Xenopoulos, Freeman, and Silva 2022) and Valorant (DeRover 2021), both
of which are round-based like CoD SnD. However, considering
Counter-Strike and Valorant specifically, research on round-level trends
seems non-existant, perhaps for one of many reasons:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Both have economic aspects that can create clear advantages for one
  side in a given round, given how prior rounds played out. CoD has no
  such equivalent, except for perhaps ``kill streaks'', which
  infrequently play a role.
\item
  Teams play either offense or defense for many consecutive rounds. (The
  former has a 14-14-1 format and the latter uses a 12-12-1 format.) On
  the other hand, teams in CoD SnD rotate roles every round, analogous
  to a 1-1-1-1-1-1-1 format for home advantage in best-of-seven series
  for professional sports like the MLB, NBA, and NHL.\footnote{1-1-1-1-1-1-1
    is not used today in these leagues, but it was at least once in each
    league.} While theoretically one might be able to account for any
  kind of format, such as a 5-5-1, the rotation of team sides every
  round is convenient for convincing ourselves that rounds could
  reasonably be modeled as i.i.d. Bernoulli trials.
\item
  Both Counter-Strike and Valorant have overtime rules---team must win
  by two rounds---which can make end-of-series sequences difficult to
  model. CoD SnD does not have overtime rules.
\end{enumerate}

To our knowledge, there is no existing public statistical research on
CoD SnD format, beyond descriptive analysis on social media. While
intra-round trends may be more directly applicable to teams looking for
an advantage on their competition, broader investigation on a common
concept like the hot-hand phenomenon in a new field---particularly one
that is less subjected to factors that may be difficult to control for,
e.g.~weather---should be useful as a precedent or supplement for future
researchers.

@@

In a best-of-\(s\) format, assuming a constant round win probability
\(\phi\) (\(0 \leq \phi \leq 1\)) for one team\footnote{If
  \(\phi > 0.5\), we can say that this is the stronger team.}, the
probability that a series ends in \(r\) rounds (\(r \leq s\)),
\(\hat{\Phi}(r)\) is

\begin{equation}\label{eq:m}
m = \frac{s + 1}{2}
\end{equation}

\begin{equation}\label{eq:series-length}
\hat{\Phi}(r) = \frac{(r - 1)!}{(m - 1)!(s - r)!}(\phi^{m}(1 - \phi)^{r - m} + \phi^{r - m}(1 - \phi)^m)
\end{equation}

\(s = 11\) rounds for CoD SnD.\footnote{\(\phi\) and other symbols are
  selected so as to reserve symbols like \(p\) for usage in other
  contexts without causing confusion for the reader. Upper-case symbol,
  e.g.~\(\Phi\), are consistently used in this paper to represent
  proportions, while lower-case symbols are used to represent Bernoulli
  trials success probability. ``Hats'', e.g.~\(\hat{\Phi}\), are used to
  convey expectations, while bare symbols symbols convey observational
  data.} For example, assuming \(\phi = 0.5\), the probability of a
series ending in 9 rounds in CoD is

@@

\begin{equation}\label{eq:chi-squ}
\chi^2 = \sum^{r \in R} \frac{(\Phi(r) - \hat{\Phi}(r))^2}{\hat{\Phi}(r)}.
\end{equation}

@@

\begin{equation}\label{eq:pwr}
p_0(\text{win} | r) = p^{+|r}_0 = \begin{cases} 
\frac{m}{r} & \text{team wins series} \\ 
\frac{r - m}{r} & \text{team loses series}  
\end{cases}
\end{equation}

@@

As shown by MS with their Theorem 1, we should expect the proportion of
rounds wins immediately following a streak of \(k\) rounds wins for a
series lasting \(r\) rounds, \(\hat{P}^{+|k,r}_{MS}\), to be strictly
less than \(p^{+|r}\).\footnote{See Appendix E.1 of MS for the proof.}
Unfortunately, there does not exist a closed form representation for the
expected value of \(\hat{P}^{+|k,r}_{MS}\) for \(k > 1\). Nonetheless,
one may recursively build a collection of dictionaries to compute the
expectation analytically, or, more conveniently, one may run simulations
to estimate the expected value, as We choose to do.\footnote{We have
  adapted the code from Vafa (2017), which implements MS's framework
  (2018).}

Although it would be convenient to emulate the hypothesis testing for
difference of proportions performed by MS (and GVT), our context is
fundamentally different from that of MS, who focus on longitudinal data
in controlled settings.\footnote{GVT also performs statistical tests on
  shots from players in live games, i.e.~``observational'' data, but
  they note that their findings are likely affected player shot
  selection in the face of defensive strategy by the opposing team.} The
number of trials is fixed in their experimental designs, but in CoD SnD,
the number of rounds played is determined as a function of the max
number of possible rounds (\(s\)) and whether or not the team wins the
series. The Bernoulli trial success probability, i.e.~a round win in CoD
SnD, is not independent of the opponent. Consequently, a statistical
test of the difference in \(\hat{P}^{+|k,r}\) and \(\hat{P}^{-|kr}\), as
performed by MS to evaluate their hypothesis regarding post-streak
success rates, is not completely appropriate, although useful as a
reference.

@@

\begin{equation}\label{eq:pwkr}
p_0(\text{win} | k, r) = p^{+|k,r}_0 = \begin{cases}
\frac{m - k}{r - k} & \text{team wins series} \\
\frac{s - m - k}{r - k} & \text{team loses series}
\end{cases}
\end{equation}

@@

We can perform a binomial test to evaluate the null hypothesis
\(\omega = \omega_0\) for the observed probability of success \(\omega\)
and a user-specified \(\omega_0\), where
\(0 \leq \omega_0 \leq 1\).\footnote{\(\omega\) and other symbols are
  chosen so as to not conflict with symbols already used elsewhere for
  in other equations, either generic or for CoD SnD specifically.} If
there are \(r^+\) observed successes in a sample of \(r\) trials and we
expect that there should be \(r * \omega_0\), the probability of
arriving at this expected number of successes is

@@

Treating the notional proportion \(p^{+|k,r}_0\) as the null
\(\omega_0\) and plugging in the observed proportion \(P^{+|k,r}\) for
\(\omega\) in Equation \ref{eq:binom} (treating proportions as
probabilities), we can evaluate the null hypothesis that the observed
probability is equal to the notional probability. If we can reject this
null hypothesis, then we can consider team momentum, represented by
post-streak success, plausible.

One can perform the same binomial test for the MS's
streak-selection-adjusted proportion, \(\hat{P}^{+|k,r}_{MS}\). FOOTY:
Given the caveats mentioned before, the results should be heeded with
caution.

We can further decompose Equation \ref{eq:pwkr} by the round \(i\)
(\(i \leq r\)) in which the streak of length \(k\) carries into.

\begin{equation}\protect\hypertarget{eq:pwkri}{}{
p_0(\text{win} | k, r, i) = p^{+|k,r,i}_0 =\{
\begin{array}{cl}
\frac{m - k}{r - k}, & \text{team wins series}, i \neq r \\
1, & \text{team wins series}, i = r \\
\frac{s - m - k}{r - k}, & \text{team loses series}, i \neq r \\
0, & \text{team loses series}, i = r
\end{array}
}
\label{eq:pwkri}
\end{equation}

Again, one might apply a binomial test to evaluate the hypothesis that
the expected proportions, both \(\hat{P}^{+|k,r,i}_0\) and
\(\hat{P}^{+|k,r,i}_{MS}\), are equal to the observed proportion
\(P^{+|k,r}\).

@@

Stepping back from the one-sided perspective of a single team's round
win probability when streaking, one can attempt to detect the hot hand
phenomenon with the Wald-Wolfowitz runs test. Under the null hypothesis,
the number of runs in a sequence of \(r\) trials, \(\zeta(r)\), is a
random variable that can take on values \(r^+\) and \(r^-\) for success
and failure respectively, with the following mean \(\mu\) and variance
\(\sigma^2\).

\begin{equation}\protect\hypertarget{eq:ww}{}{
\mu = \frac{2r^{+}r^{-}}{r} + 1, \sigma^2 = \frac{(\mu-1)(\mu-2)}{r-1}
}\label{eq:ww}\end{equation}

@@

One can subset the observed series sequences to those that violate the
run test null hypothesis and perform a test of proportions, where the
null is that the observed rate \(\psi^+\) of such series is equal to the
frequency rate \(\psi_0\). The test statistic \(Z\)

>>>>

\title{The Hot Hand Fallacy in Call of Duty}

@@

\begin{abstract}
Our research investigates patterns in round win percentages in
professional Search and Destroy (SnD) matches of the popular
first-person shooter game Call of Duty (CoD). First, we find evidence
suppOrting the hypothesis that round win probability can be modeled as a
constant across rounds in the series, although not at the naive 50\%.
Second, we examine post-streak round win probability, given the series
outcome. Given streak length, series length, and series winner, we find
no evidence that post-streak win rate is significantly different than
expected, suggesting that the perception of momentum at the team level
is deceiving. Wald-Wolfowitz run tests also fail to provide evidence for
the hot hand phenomenon.
\end{abstract}

@@

Call of Duty (CoD), first released in 2003, is one of the most popular
first-person shooter (FPS) video game franchises of all-time. The most
popular mode in the competitive scene is ``Search and Destroy'' (SnD),
in which one team tries to destroy one of two designated bomb sites on
the map, defended by the opposing team.\footnote{SnD bears resemblance
  to ``Bomb Defusal'' in Counter-Strike and ``Plant/Defuse'' in
  Valorant, two other FPS games played in more popular professional
  leagues.}

In professional CoD SnD, the two teams\footnote{In the 2020 season,
  teams played with five players each; in the 2021 and 2022 seasons,
  teams played with four players each.} take turns playing offense and
defense every round. The first to six round wins (best-of-11 format) is
declared the series winner.\footnote{Each round has a two-minute time
  limit, not counting the potential extension granted if a bomb is
  planted.} A round can end in one of four ways:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  One team eliminates all members of the other team prior to a bomb
  plant. (Eliminating team wins.)
\item
  The offensive team eliminates all members of the defensive team after
  a bomb plant.\footnote{\begin{itemize}
    \tightlist
    \item
      The bomb can be picked up by any member of the offensive team.
    \item
      The bomb carrier is not obstructed by carrying the bomb
      (i.e.~movement is the same, weapon usage is the same, etc.).
    \item
      The defense does not get any visual indication for who is carrying
      the bomb.
    \item
      A bomb plant takes five seconds. The timer resets if the player
      stops planting site prior to completing it.
    \item
      A bomb defuse takes seven seconds. The timer resets if the player
      ``drops'' the bomb.
    \item
      Once the bomb is planted, the round timer is reset to 45 seconds.
    \end{itemize}} (Offense wins.)
\item
  The defense defuses the bomb after a bomb plant.\footnote{Often the
    defensive team will try to eliminate all team members prior to
    making the defuse, but in some cases, they may try to ``ninja''
    defuse.} (Defense wins.)
\item
  The offense does not make a plant by the time the round timer ends.
  (Defense wins.)
\end{enumerate}

We adopt the terminology ``series'' to refer to a single best-of-11
matchup, so as to mirror the terminology of playoff series in
professional leagues like the National Basketball Association (NBA),
National Hockey League (NHL), and Major League Baseball (MLB). A
``game'' or a ``match'' in such leagues is analogous to a ``round'' of
CoD SnD for the purposes of this paper.\footnote{Further, these are
  analogous to a trial in statistical analysis of Bernoulli events.}

SnD is not the only game type played in competitive CoD. Teams play in a
head-to-head, best-of-five format, where SnD is always played as the
second and fifth game ``maps'', as most players refer to them (if a
fifth map is needed). The best-of-five matchup could also be called a
``series'', but since we analyze only the SnD game types, we refer to
SnD games as series.

@@

CoD has roughly gone through three eras of professional gaming: (1)
Major League Gaming (MLG) tournaments prior to 2016; (2) the CoD World
League (CWL), initiated in 2016; and (3) the 12-franchise CoD League
(CDL), operating since 2020. The CDL has completed three year-long
``seasons'' as of August 2022.\footnote{CoD is fairly unique compared to
  other esports in that it runs on an annual lifecycle, with releases
  coming in the late fall. A new game under the same brand---Call of
  Duty---is published every year by a rotating set of developers. Each
  new game bears resemblance to past ones, often introducing relatively
  small variations (``improvements'') to graphics, game modes, and other
  facets of gameplay. During the CDL era, the games released have been
  Modern Warfare (2020), Cold War (2021) and Vanguard (2022).}

The data set consists of all SnD matches played in major tournaments and
qualifiers during the CDL era, totaling 7,792 rounds across 852
series.\footnote{288 of the series occur in major tournaments, which are
  considered to be more ``competitive'' than the qualifiers since they
  are played in person (COVID-permitting), whereas the qualifiers are
  played online.} Data has been collected in spreadsheets by community
member ``IOUTurtle''.\footnote{Raw data:
  \url{https://linktr.ee/CDLArchive}. Author:
  \url{https://twitter.com/IOUTurtle}. Processed data can be found at
  \url{https://github.com/tonyelhabr/fps_round_games/blob/master/data/cod_rounds.csv}.}

6-0 sweeps make up 4.7\% of all series, while 6-5 series make up 23.7\%
of series. The observed offensive round win percentage across all rounds
is \(\tau_O\) = 47.8\%.\footnote{Offensive round win percentage has been
  nearly constant across the three games during the CDL era: 1. 47.2\%
  in MW (2020) 2. 47.9\% in Cold War (2021) 3. 48.1\% in Vanguard (2022)}
Table \ref{tbl:cod-o-win-prop-by-series-state} shows round win
percentages by series ``state'' (i.e.~the number of round wins by each
team prior to an upcoming round). Offensive round win rate is not quite
constant, although never veers more than 10\% from \(\tau_O\).

@@

At tournaments, higher-seeded (``better'') teams choose whether they
want to start on offense in the SnD series played as the second or fifth
map, and their opponent is assigned to start on offense in the SnD map
not chosen.\footnote{Hardpoint has been played as the first and fourth
  maps in a matchup. The variant for the third map was Domination in the
  2020 season, and Control in the 2021 and 2022 seasons.}\footnote{Anecdotally,
  the higher-seeded teams tend to choose to play defense in the first
  round of the SnD series played as the second game, although this
  choice is not consistent across teams, or with the same team over
  time.} Outside of tournaments, teams all play each other twice
throughout the season, constituting qualifiers. Each team gets to play
the higher-seeded role once in their head-to-head matchups with a given
team, despite their win-loss records. Overall, we can argue that
potential bias due to stronger teams playing defense---or, if offense,
if offense happens to be generally advantageous for a given map---is
very minimal.

@@

There have been a handful of studies of the distribution of games played
in a series of a professional sport. Most assume a constant probability
\(\phi\) of a given team winning a game in the series, regardless of the
series state. Mosteller (1952) observed that the American League had
dominated the National League in MLB World Series matchups through 1952,
implying that games should not be modeled with a constant (i.e.~the
null) \(\phi_0 = 0.5\). Mosteller proposed three approaches for
identifying the optimal constant probability value of the stronger team
in the World Series, finding \(\phi \approx 0.65\) in each case: (1)
solving for \(\phi\) from the observed average number of games won by
the loser of the series, which he called the ``method of moments''
approach; (2) maximizing the likelihood that the sample would have been
drawn from a population in which the probability of a team winning a
game is constant across the series (i.e.~maximum likelihood), and (3)
minimizing the chi-squared goodness of fit statistic \(\chi^2\) as a
function of \(\phi\).

Chance (2020) re-examines the constant probability notion in the MLB
World Series (1923--2018), the NBA Finals (1951--2018), and the NHL
Stanley Cup (1939--2018). Chance found strong evidence against the null
hypothesis of \(\phi_0 = 0.5\) in the MLB and NHL championship series
when applying Mosteller's first and second methods.

Momentum, one of most discussed topics in sports analytics, goes
hand-in-hand with a discussion of the nature of series
outcomes.\footnote{We often use ``streaks'' and momentum
  interchangeably, but as Steeger et al.~((2021)) note, momentum implies
  dependence between events, whereas streaking does not.} Two opposing
fallacies are observed in the context of momentum: the ``gambler's
fallacy'' (negative recency) and the ``hot hand fallacy'' (positive
recency). Per Ayton et al. (2004), negative recency is ``the belief
that, for random events, runs of a particular outcome \ldots{} will be
balanced by a tendency for the opposite outcome'', while positive
recency is the expectation of observing future results that match recent
results.

Studying both player streaks and team streaks in basketball, in both
observational and controlled settings. Gilovich et al. (1985),
henceforth GVT, ido not find evidence for the hot hand phenomenon.
However, Miller and Sanjurjo (2018), henceforth MS, provided a framework
for quantifying streak selection bias, which effectively works in the
manner posited by the gambler's fallacy. Specifically, MS say that a
``bias exists in a common measure of the conditional dependence of
present outcomes on streaks of past outcomes in sequential data''
implying that, under i.i.d. conditions, ``the proportion of successes
among the outcomes that immediately follow a streak of consecutive
successes is expected to be strictly less than the underlying
(conditional) probability of success''. When applying streak selection
bias to GVT's data, MS came to the opposite conclusions as GVT.

Other research has borrowed a common approach from the field of quality
control, looking at unlikely sequences of events with the Wald-Wolfowitz
runs test (Peel and Clauset 2015, Steeger et al.~2021). Peel and Clauset
found no evidence for unlikely sequences of scoring events in the NHL,
College Football (CFB), and National Football League (NFL), but did in
the NBA. As a check on their entropy approach to momentum
identification, Steeger et al.~found several NHL teams with sequence of
wins in the 2018-2019 regular season violating the Wald-Wolfowitz null
hypothesis.

@@

Despite the plethora of existing research on games played in a series
and momentum in sports, these topics have yet to be investigated
heavily, if at all, in esports. Work has been done to examine
intra-round win probability in other FPS titles such as Counter-Strike
(Xenopoulos, Freeman, and Silva 2022) and Valorant (DeRover 2021), both
of which are round-based like CoD SnD. However, considering
Counter-Strike and Valorant specifically, research on round-level trends
seems non-existent, perhaps for one of several reasons:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Both have economic aspects that can create clear advantages for one
  team in a round, given how prior rounds played out. CoD has no such
  equivalent, except for perhaps ``kill streaks'', which infrequently
  occur.
\item
  Teams play either offense or defense for many consecutive rounds. (The
  former has a 15-15-1 format and the latter uses a 12-12-1 format.) On
  the other hand, teams in CoD SnD rotate roles every round, analogous
  to a 1-1-1-1-1-1-1 format for home advantage in best-of-seven series
  for professional sports like the MLB, NBA, and NHL.\footnote{1-1-1-1-1-1-1
    is not used today in these leagues, but it was at least once in each
    league.} While theoretically one might be able to account for any
  kind of format, such as a 5-5-1, the rotation of team sides every
  round is convenient for convincing ourselves that rounds could
  reasonably be modeled as i.i.d. Bernoulli trials.
\item
  Both Counter-Strike and Valorant have overtime rules---team must win
  by two rounds---which can make end-of-series sequences difficult to
  model. CoD SnD does not have overtime rules.
\end{enumerate}

To our knowledge, there is no existing public statistical research on
the CoD SnD format, beyond descriptive analysis on social media. While
intra-round trends may be more directly applicable to teams looking for
an advantage on their competition, broader investigation of a concept
like the hot-hand fallacy in a different sport---particularly one that
is less subjected to factors that may be difficult to control for,
e.g.~weather---should be useful as a precedent or supplement for future
researchers.

@@

In a best-of-\(s\) format, assuming a constant round win probability
\(\phi\)\footnote{If \(\phi > 0.5\), we can say that this is the
  stronger team.} (where \(0 \leq \phi \leq 1\)) for one
team\footnote{\(\phi\) and other symbols are selected so as to reserve
  symbols like \(p\) for usage in other contexts without causing
  confusion for the reader. Upper-case symbol, e.g.~\(\Phi\), are
  consistently used in this paper to represent proportions, while
  lower-case symbols corresponding to upper-case symbols, e.g.~\(\phi\),
  are used to represent Bernoulli trial success rates. ``Hats'',
  e.g.~\(\hat{\Phi}\), are used to convey expectations, while bare
  symbols symbols convey observational data.}, the expected proportion
of series ending in \(r\) rounds (\(r \leq s\)) is given by Equation
\ref{eq:series-length}.

\begin{equation}\label{eq:m}
m = \frac{s + 1}{2}.
\end{equation}

\begin{equation}\label{eq:series-length}
\hat{\Phi}(r) = \frac{(r - 1)!}{(m - 1)!(s - r)!}(\phi^{m}(1 - \phi)^{r - m} + \phi^{r - m}(1 - \phi)^m).
\end{equation}

For example, assuming \(\phi = 0.5\), the probability of a series ending
in nine rounds in CoD SnD, where \(s=11\), is

@@

\begin{equation}\label{eq:chi-squ}
\chi^2 = \sum^{r \in R} \frac{(\Phi(r) - \hat{\Phi}(r))^2}{\hat{\Phi}(r)},
\end{equation}

@@

\begin{equation}\label{eq:pwr}
p_0(\text{win} | r) = p^{+|r}_0 = \begin{cases} 
\frac{m}{r} & \text{team wins series}, \\ 
\frac{r - m}{r} & \text{team loses series},
\end{cases}
\end{equation}

@@

As shown by MS with their Theorem 1, we should expect the proportion of
round wins immediately following a streak of \(k\) rounds wins for a
series lasting \(r\) rounds, \(\hat{P}^{+|k,r}_{MS}\), to be strictly
less than \(p^{+|r}\).\footnote{See Appendix E.1 of MS for the proof.}
Unfortunately, there does not exist a closed form representation for the
expected value of \(\hat{P}^{+|k,r}_{MS}\) for \(k > 1\). Nonetheless,
one may run simulations to estimate the expected value, as we choose to
do.\footnote{We have adapted the code from Vafa (2017), which implements
  MS's framework (2018).}

Although one might be tempted to mirror the hypothesis testing for
difference of proportions performed by MS (and GVT), our context is
fundamentally different from that of MS, who focus on longitudinal data
in controlled settings.\footnote{GVT also perform statistical tests on
  shots from players in live games, i.e.~``observational'' data, but
  they note that their findings are likely affected player shot
  selection in the face of defensive strategy by the opposing team.} The
number of trials is fixed in their experimental designs, but in CoD SnD,
the number of rounds played is determined as a function of the max
number of possible rounds (\(s\)) and whether or not the team wins the
series. The Bernoulli trial success probability, i.e.~the single round
win rate in CoD SnD, is not independent of the opponent. Consequently, a
statistical test of the difference in \(\hat{P}^{+|k,r}\) and
\(\hat{P}^{-|kr}\), as performed by MS to evaluate their hypothesis
regarding post-streak success rate, is not completely appropriate,
although useful as a reference.

@@

\begin{equation}\label{eq:pwkr}
p_0(\text{win} | k, r) = p^{+|k,r}_0 = \begin{cases}
\frac{m - k}{r - k} & \text{team wins series}, \\
\frac{s - m - k}{r - k} & \text{team loses series},
\end{cases}
\end{equation}

@@

We can perform a binomial test to evaluate the null hypothesis
\(\omega = \omega_0\) for the observed probability of success \(\omega\)
and a user-specified \(\omega_0\) (where \(0 \leq \omega_0 \leq 1\)). If
there are \(r^+\) observed successes in a sample of \(r\) trials and we
expect that there should be \(r * \omega_0\), the probability of
arriving at this expected number of successes is

@@

Treating the notional proportion \(\hat{P}^{+|k,r}_0\) as the null
\(\omega_0\) and plugging in the observed proportion \(P^{+|k,r}\) for
\(\omega\) in Equation \ref{eq:binom} (treating proportions as
probabilities), we can evaluate the null hypothesis that the observed
probability is equal to the notional probability. If we can reject this
null hypothesis, then we can consider team momentum, represented by
post-streak success, plausible.

One can perform the same binomial test for the MS's
streak-selection-adjusted proportion,
\(\hat{P}^{+|k,r}_{MS}\).\footnote{Given the caveats mentioned before,
  these results should be heeded with caution.}

We can further decompose Equation \ref{eq:pwkr} by the round \(i\)
(where \(i \leq r\)) in which the streak of length \(k\) carries into.

\begin{equation}\label{eq:pwkri}
p_0(\text{win} | k, r, i) = p^{+|k,r,i}_0 = \begin{cases}
\frac{m - k}{r - k}, & \text{team wins series}, i \neq r, \\
1, & \text{team wins series}, i = r, \\
\frac{s - m - k}{r - k}, & \text{team loses series}, i \neq r, \\
0, & \text{team loses series}, i = r.
\end{cases}
\end{equation}

Again, we caapply a binomial test to evaluate the hypothesis that the
expected proportions, \(\hat{P}^{+|k,r,i}_0\) and
\(\hat{P}^{+|k,r,i}_{MS}\) separately, are equal to the observed
proportion \(P^{+|k,r}\).

@@

Stepping back from the one-sided perspective of a single team's round
win probability when streaking, one can attempt to detect the hot hand
phenomenon with a Wald-Wolfowitz runs test. Under the null hypothesis,
the number of runs in a sequence of \(r\) trials, \(\zeta(r)\), is a
random variable that can take on values \(r^+\) and \(r^-\) for success
and failure respectively, with the following mean \(\mu\) and variance
\(\sigma^2\).

\begin{equation}\label{eq:ww}
\mu = \frac{2r^{+}r^{-}}{r} + 1, \sigma^2 = \frac{(\mu-1)(\mu-2)}{r-1}
\end{equation}

@@

One can subset the observed series sequences to those that violate the
null hypothesis for the run test and perform a test of equal
proportions, where the null is that the observed rate of a sequence,
\(\psi\), is equal to the expected rate of the sequence, \(\hat{psi}\).
The test statistic is
